|

### 2020-05-18 paper/Manuscript.Rmd
Incorporate changes

# Introduction
<!--
Barbara:
A trick I learned in collaborative writing with git is to make sure each new sentence is on a new line.
This way, you can effectively use a git diff to see changes in individual sentences instead of having to scour through an entire paragraph.
I'll add this, but only in the paragraphs I've changed, so you do not get inundated with differences!
-->
The (social) sciences are amidst a paradigm shift towards open science. 
Several immediate causes stirred up support for this transition; including several highly publicized cases of scientific fraud [as discussed by @leveltFailingScienceFraudulent2012], increasing awareness of questionable research practices [@johnMeasuringPrevalenceQuestionable2012], and the replication crisis [@shroutPsychologyScienceKnowledge2018].
However, open science is not merely a cure for this crisis - it is also an opportunity [@adolphOpenBehavioralScience2012].
Technological advances enable researchers to more easily conduct reliable, cumulative, collaborative science [@nosekScientificUtopiaOpening2012].
Capitalizing on these advances has the potential to accellerate scientific progress [see also @coyneReplicationInitiativesWill2016].

@@

<!-- LB: Although Open Science is advocated by many, there is no clear definition of Open Science (ref multiple definitions). An influential operationalisation of Open Science has been de TOP guidlines. -->
One of the most impactful contributions to open science has been the development of the TOP-guidelines [@nosekPromotingOpenResearch2015a], and the effort to urge publishers to adopt them. These guidelines describe eight standards for open science: 1) Comprehensive citation of literature, data, materials, and methods; 2) sharing data, 3) sharing the code required to reproduce analyses, 4) sharing new research materials, and 5) sharing details of the design and analysis; 6) pre-registration of studies before data collection, and 7) pre-registration of the analysis plan prior to analysis; and 8) replication of published results. WORCS defines the goals of open science in terms of these guidelines, and is designed to facilitate meeting each of these guidelines, with one exception: We do not address replication of published results, because replication relates to the subject of a research project, not to its execution. We group the remaining seven guidelines into three categories: Citation (1), sharing (2-5), and preregistration (6-7).

@@

Data sharing is important for computational reproducibility and secondary analysis. Computational reproducibility means that a third party can exactly recreate the results from the raw data, using the published analysis code. Secondary analysis means that a third party can conduct sensitivity analyses, explore alternative explanations, or even use existing data to answer a different research question. From an open science perspective, data sharing is always desirable. From a practical point of view, it is not always possible. Data sharing may be impeded by legal constraints, ethical restrictions, absence of informed consent for data sharing, and privacy concerns. For example, the GDPR prohibits storing 'personal data' (information which can identify a natural person whether directly or indirectly) on a server outside the EU. Different rules may apply to anonymized or pseudonimized data. Before sharing any data, it is recommended to obtain approval from an internal (ethical) review board, guidance from Research Data Management Support, and informed consent from participants.

<!-- LB: As I reader, I would prefer to hear about data sharing options before the synthetic data part (which is only relevant to those who cannot share their data). E.g. can I indicate a point in my workflow (e.g. after removing some privacy sensitive columns of my data) to share the as-raw-as-possible data ?-->When data can be shared, it is important to version control the (anonymized/pseudonymized) raw data upon collection, and avoid changing it. Any data processing steps should be documented in the code, instead of applied to the raw data. This way, the code will be a pipeline from the raw data to the published results.

When data cannot be shared, researchers should aim to safeguard the potential for computational reproducibility and secondary analysis as much as possible.
WORCS uses two solutions to accomplish this goal.
The first solution is publishing a *checksum* of the raw data upon collection.
A checksum is the result of an algorithm applied to a file, and represents the contents of this file in 32-characters.
Any change to the underlying file results in a different checksum. Thus, one can use a checksum to verify that the contents of a file remain unchanged.
If researchers publish the checksum of their raw data, in addition to the complete analysis code, then they can prove to an independent auditer that running the public analysis code on the private dataset with the correct checksum results in the published results of their work.

The second solution is to share a synthetic dataset with similar characteristics to the real data. Synthetic data mimic the level of measurement and (conditional) distributions of the real data. Sharing synthetic data allows any third party to 1) verify that the published code runs, 2) debug the code, and 3) write valid code for alternative analyses. However, it is important to know that complex multivariate relationships present in the real data are often lost in synthetic data. Thus, findings from the real data might not be replicated in the synthetic data, and findings in the synthetic data *should not be substantively interpreted*. Still, sharing synthetic data facilitates data requests from third parties. A third party can write analysis code based on the synthetic data, and send it to the authors who evaluate it on the real data and send back the results. The ["Pull request"](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request) functionality on GitHub is a particularly user-friendly way to do this.
<!--Barbara: I'm not sure this is the case. Pull requests are useful if someone wants to change or add to a project, but this would not involve a change to an existing project, but rather a new project/analysis with "old" data. I would remove this suggestion altogether.-->

The procedure for data sharing in WORCS is as follows. First, researchers load the data into R. This process should be documented in the file `prepare_data.R`, which is automatically created by the WORCS-template. This file is intended to document any steps that must be taken to get your data *into* R, and that are not intended to be repeated when re-running the analyses. Use this file to report any preprocessing required to obtain a neat `data.frame`. For example, if the data was originally in SPSS format, this file might just contain the code required to read the SPSS file.

Then, researchers must decide whether to use open or closed data, and call one of these convenience functions:

* `open_data(data)`: This function stores the data object (a `data.frame` or `matrix`) in a spreadsheet file called `data.csv`, stores a checksum for the file, and creates a codebook. All of these files should now be (manually) added to the 'Git' repository, committed, and pushed to GitHub. Once the GitHub repository is made public, these data will be open to the public.
* `closed_data(data)`: This function stores the data in a local file called `data.csv`, which is added to the `.gitignore` file so it cannot be accidentally added to the 'Git' repository. The function also creates a synthetic dataset called `synthetic_data.csv`. The user should commit all changed files and push them to GitHub. Once the GitHub repository is made public, people will have access to a checksum for the raw data that exist on your local machine, a codebook describing those data, and a synthetic copy of the data.

<!--Barbara: Is it indeed the case that data by default is loaded as a .csv? How about (raw) text data; can worcs accommodate a .txt file, for instance (in case of text analysis)? Or should this all be done in prepare_data.R?-->

Some users may intend to share their data through a platform with restricted access (like an institutional repository), but not publically through GitHub. In this case, it is recommended to use `closed_data()`, and to manually upload the original data `data.csv` to the restricted access platform. If users wish to share their data through the Open Science Framework, it is sufficient to connect the OSF page to the GitHub repository.

The `prepare_data.R` file describes the data wrangling process to derive a spreadsheet ready for analysis. The final line of this file will often be `open_data()` or `closed_data()`. The code for analysis of the data, on the other hand, will typically begin with `load_data()`. This function loads the real data if it is available (i.e., the file `data.csv` exists, and the checksum is correct), and otherwise, loads the synthetic data. This will have the effect that third party users who copy the GitHub repository will automatically load the synthetic dataset, whereas the study authors, who have the real data stored locally, will automatically load the real data.

@@

4. Optional: Add Materials to repository. It is possible to solicit feedback (through issue/pull request) and later acknowledge outside contributions
4. Collect data and either add to repository, or register hash of raw datafile
5. Conduct analyses and write; commit regularly (ideally, after completing a small and clearly defined task), with informative commit messages. Call renv::snapshot() to save the state of the project library to the lockfile (called renv.lock),
6. Create OSF project, connect to GitHub, mention OSF page in manuscript
7. Optional: Publish preprint in a not-for-profit preprint repository (check Sherpa Romeo) http://sherpa.ac.uk/romeo/index.php
7. Submit
8. Make repository public -->

>>>>

# Introduction

The (social) sciences are amidst a paradigm shift towards open science. 
Several immediate causes stirred up support for this transition; including several highly publicized cases of scientific fraud [as discussed by @leveltFailingScienceFraudulent2012], increasing awareness of questionable research practices [@johnMeasuringPrevalenceQuestionable2012], and the replication crisis [@shroutPsychologyScienceKnowledge2018].
However, open science is not merely a cure for this crisis - it is also an opportunity [@adolphOpenBehavioralScience2012].
Technological advances enable researchers to more easily conduct reliable, cumulative, collaborative science [@nosekScientificUtopiaOpening2012].
Capitalizing on these advances has the potential to accellerate scientific progress [see also @coyneReplicationInitiativesWill2016].

@@

Although Open Science is advocated by many, there is no clear definition of Open Science (LOEK? ref multiple definitions).
The TOP-guidelines have been influential as a concrete operationalisation of Open Science principles [@nosekPromotingOpenResearch2015a].
These guidelines describe eight standards for open science: 1) Comprehensive citation of literature, data, materials, and methods; 2) sharing data, 3) sharing the code required to reproduce analyses, 4) sharing new research materials, and 5) sharing details of the design and analysis; 6) pre-registration of studies before data collection, and 7) pre-registration of the analysis plan prior to analysis; and 8) replication of published results. WORCS defines the goals of open science in terms of these guidelines, and is designed to facilitate meeting each of these guidelines, with one exception: We do not address replication of published results, because replication relates to the subject of a research project, not to its execution. We group the remaining seven guidelines into three categories: Citation (1), sharing (2-5), and preregistration (6-7).

@@

Data sharing is important for computational reproducibility and secondary analysis. Computational reproducibility means that a third party can exactly recreate the results from the original data, using the published analysis code. Secondary analysis means that a third party can conduct sensitivity analyses, explore alternative explanations, or even use existing data to answer a different research question. From an open science perspective, data sharing is always desirable. From a practical point of view, it is not always possible. Data sharing may be impeded by legal constraints, ethical restrictions, absence of informed consent for data sharing, and privacy concerns. For example, one concern that applies when working with human participant data, is pseudonimization. It is crucial to pseudonimize data as soon as they are collected, by removing or deleting any sensitive personal information and contact details. Pseudononimized data can often be shared if participants have provided informed consent to that effect. Nevertheles, it is important to note that the European GDPR prohibits storing 'personal data' (information which can identify a natural person whether directly or indirectly) on a server outside the EU, unless it offers an "adequate level of protection". Although different rules may apply to pseudonimized data, there are many repositories that are GDPR compliant, such as the European servers of the Open Science Framework. Different Universities, countries, and funding bodies also have their own repositories that are complient with local legistation. Before sharing any human participant data, it is recommended to obtain approval from an internal (ethical) review board, guidance from Research Data Management Support, and informed consent from participants.

<!-- LB: As I reader, I would prefer to hear about data sharing options before the synthetic data part (which is only relevant to those who cannot share their data). E.g. can I indicate a point in my workflow (e.g. after removing some privacy sensitive columns of my data) to share the as-raw-as-possible data ?-->
If data cannot be shared, researchers should aim to safeguard the potential for computational reproducibility and secondary analysis as much as possible. The `worcs` package uses two solutions to accomplish this goal.
The first solution is to publish a *checksum* of the original data file.
Think of a checksum as a 32-character summary of the contents of a file, or alternatively, a "one word" description of the file.
Any change to the file will result in a different checksum. 
Thus, one can use a checksum to verify the identity of a file, and ensure that its contents are unchanged.
When data cannot be shared, the risk of fraud or abuse can be mitigated by publishing a checksum for the original data, in addition to the complete analysis code.
Using the checksum to verify the identity of a private dataset, researchers can prove to an independent auditer that running the public analysis code on the private data results in the published results of their work.

The second solution implemented in `worcs` is to share a synthetic dataset with similar characteristics to the real data. Synthetic data mimic the level of measurement and (conditional) distributions of the real data [see @nowokSynthpopBespokeCreation2016]. Sharing synthetic data allows any third party to 1) verify that the published code works, 2) debug the code, and 3) write valid code for alternative analyses. It is important to note that complex multivariate relationships present in the real data are often lost in synthetic data. Thus, findings from the real data might not be replicated in the synthetic data, and findings in the synthetic data *should not be substantively interpreted*. Still, sharing synthetic data facilitates data requests from third parties. A third party can write analysis code based on the synthetic data, and send it to the authors who evaluate it on the real data and send back the results.

### Processing data in WORCS

Regardless of whether the data will be open or closed, it is important that the raw data will be left unchanged as much as possible. This eliminates human error. Any alterations to the data - including processing steps and even error corrections - should be documented in the code, instead of applied to the raw data. This way, the code will be a pipeline from the raw data to the published results. Every `worcs` project contains an empty R-script called `prepare_data.R`. As soon as the data are collected, researchers should use this file to document the steps necessary to load the data into R, pseudonimize it, and prepare it for analysis. For example, if the data was originally in SPSS format with IP addresses and GPS location, this file might just contain the code required to read the SPSS file, and to remove those columns of sensitive personal information. As soon as the data are pseudonimized and processed into a 'tidy' format [i.e., data that can be stored as a spreadsheet, see @wickhamTidyData2014], the researcher should version control some indelible record of the raw data, and avoid making any further changes to the version controlled file.

The `worcs` package offers two functions for version controlling a record of the data: One for open, and one for closed data:

* `open_data(data)`: This function stores the `data` object (a `data.frame` or `matrix`) in a spreadsheet file (called `data.csv` by default).
This file should now be (manually) added to the 'Git' repository, committed, and pushed to GitHub.
Once the GitHub repository is made public, these data will be open to the public.
Assume that, once you do this, the data cannot be un-shared.
* `closed_data(data)`: This function also stores the data in a local file (called `data.csv` by default), but that file is added to the `.gitignore` file so it cannot be accidentally added to the 'Git' repository.
The function also adds a checksum for the original data in the `.worcs` project file, and creates a synthetic dataset (called `synthetic_data.csv` by default).
The user should commit all changed files and push them to GitHub.
Once the GitHub repository is made public, people will have access to a checksum for the original data that exist on your local machine, a codebook describing those data, and a synthetic copy of the data.

The final line of the `prepare_data.R` file should thus be either `open_data()` or `closed_data()`. The analysis code, on the other hand, should begin with `load_data()`. This function loads the real data if it is available on the user's computer, and otherwise, loads the synthetic data. This will have the effect that third party users who copy the GitHub repository will automatically load the synthetic dataset, whereas the study authors, who have the real data stored locally, will automatically load the real data. This function makes it possible for reviewers, coauthors, and auditors, to write analysis scripts without requiring access to the original data. They can simply start the script with `load_data()`, and write their code based on the synthetic data. Then, they can submit their code to you - by email or as a ["Pull request"](https://help.github.com/en/github/collaborating-with-issues-and-pull-requests/creating-a-pull-request) on GitHub. When you run their code on your system, `load_data()` will detect the original data, and use that to run their script. You can then return the results to the third party.

As of version 0.1.1, the `worcs` package only stores data as `.csv` files. As explained above, this is intentional, because `.csv` files are human- and machine readable, and because data typically need to be in a spreadsheet format for analysis. Many types of data can be represented as a spreadsheet - text corpora can have one row per document, and EEG- or EKG waveforms can have one row per measurement occasion. The `prepare_data.R` file should document any steps required to convert these data into tabular format. If it is data must be stored in a different file type anyway, readers are encouraged to follow the development of the `repro` package [@repro], which will enable such enhancements.

Some users may intend to make their data openly available through a restricted access platform, such as an institutional repository, but not publically through GitHub. In this case, it is recommended to use `closed_data()`, and to manually upload the original `data.csv` file to the restricted access platform. If users wish to share their data through the Open Science Framework, it is sufficient to connect the OSF page to the GitHub repository as an [Add-on](https://help.osf.io/hc/en-us/categories/360001550973-Add-ons).

@@

4. Optional: Add Materials to repository. It is possible to solicit feedback (through issue/pull request) and later acknowledge outside contributions
4. Collect data and either add to repository, or register hash of original datafile
5. Conduct analyses and write; commit regularly (ideally, after completing a small and clearly defined task), with informative commit messages. Call renv::snapshot() to save the state of the project library to the lockfile (called renv.lock),
6. Create OSF project, connect to GitHub, mention OSF page in manuscript
7. Optional: Publish preprint in a not-for-profit preprint repository (check Sherpa Romeo) http://sherpa.ac.uk/romeo/index.php
7. Submit
8. Make repository public -->
